---
title: "Exercises: Intro to R Markdown"
subtitle: "Tools and Workflows for Reproducible Research in the Quantitative Social Sciences"
---

```{r child = "content/config/exercises_setup.Rmd"}
```

Your task in this exercise is to create an `R Markdown` document that reports an analysis based on the (synthetic) data from the *GESIS Panel Special Survey on the Coronavirus SARS-CoV-2 Outbreak in Germany*. The output should be an `HTML` document. To come up with your own ideas for toy examples that you could include in your report, make sure to consult the [codebook for the original data](https://access.gesis.org/dbk/67378).

**Note**: We have created an exemplary `R Markdown` report that you can use as inspiration or a starting point for creating your document. You can find the `.Rmd` file as well as the `HTML` output in the `solutions` folder within the workshop materials: `Example_RMarkdown.Rmd` and `Example_RMarkdown.html`.

```{block, box.title = "Exercise 1", box.body = list(fill = "white"), box.icon = "fa-star"}
Create a new `R Markdown` document, give it a meaningful name, and save it somewhere in your project folder where you will find it. The output format that we want to use for this exercise is `HTML`.
```

```{block, opts.label = "clues", box.icon = "fa-lightbulb"}
The easiest way to create a new `R Markdown` document is through the *RStudio* GUI.
```

```{block, box.title = "Solution 1", solution = TRUE, box.icon = "fa-check"}
You can create a new `R Markdown` document in *RStudio* via *File* -> *New File* -> *R Markdown* in the menu.
```

```{block, box.title = "Exercise 2", box.body = list(fill = "white"), box.icon = "fa-star"}
Create/edit the `YAML` header of the document, so that it includes the following: a meaningful title, an author name, the current date.
```

```{block, opts.label = "clues", box.icon = "fa-lightbulb"}
You can already specify some parts of the `YAML` header through the *RStudio* GUI when creating a new document. To add the current date, you can use inline `R` code in the respective field within the `YAML` header. Remember that you need to enclose the `YAML` header in --- .
```

```{block, box.title = "Solution 2", solution = TRUE, box.icon = "fa-check"}
title: "My title"
author: "My name"
date: "`r Sys.Date()`"
output:
  html_document
```

```{block, box.title = "Exercise 3", box.body = list(fill = "white"), box.icon = "fa-star"}
Include a setup chunk in your document. Use this chunk to set the `knitr` options so that the code is displayed in the output document but messages are suppressed.

Three other things we want to do in the setup chunk:

- set the "penalty to be applied when deciding to print numeric values in fixed or exponential notation" to 10
- set the number of significant digits to print to 2
```

```{block, opts.label = "clues", box.icon = "fa-lightbulb"}
For the general options (i.e., the one not related to `knitr`), you can check `?options`.

You probably do not want to include the chunk options as code output in the resulting `HTML` document. You can achieve this by setting the chunk option `include` to `FALSE` for the setup chunk.

Of course, if you want to use additional/other packages, you can also add those to the setup chunk.
```

```{block, box.title = "Solution 3", solution = TRUE, box.icon = "fa-check"}
Chunk options: {r setup, include=FALSE}
(for the code to be included in this chunk, see the following solution box)

*NB*: In your `R Markdown` document, a code block needs to start and end with three backticks. Remember that in *RStudio* you can use the keyboard shortcut <kbd>Ctrl + Alt + I</kbd> (*Windows* & *Linux*)/<kbd>Cmd + Option + I</kbd> (*Mac*) for inserting code chunks.
```

```{r setup, eval = FALSE, box.title = "Solution 3", solution = TRUE, box.icon = "fa-check"}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)

options(scipen = 10,
        digits = 2)
```

```{block, box.title = "Exercise 4", box.body = list(fill = "white"), box.icon = "fa-star"}
Now you can start to create some content for your document. You can get as creative as you want, but we suggest that the document (at least) includes the following:

- two different levels of headings

- some basic text formatting and links

- some inline code

- multiple code chunks (with different chunk options)

- at least one table

- at least one figure

We also suggest that you loosely follow the structure of a classic research paper (Research Question, Methods, Results), but each of the sections should be very brief.
```

```{block, opts.label = "clues", box.icon = "fa-lightbulb"}
To check if everything works the way you expect, you should knit the document periodically. Also remember that we have created (and also knitted) an exemplary report which you can find in the `solutions` folder.

Please note that we use several different packages in that example document. For example, we make use of the [`tidyverse` collection of packages](https://www.tidyverse.org/) for wrangling the data. If you are not familiar with that (e.g., the use of pipes `%>%`) or prefer other packages for creating tables or plots, feel free to change anything you like in that document.
```

```{block, box.title = "Solution 4", solution = TRUE, box.icon = "fa-check"}
Feel free to have a look at `Example_RMarkdown.Rmd` and the resulting `Example_RMarkdown.html` for some examples and inspiration. Also make sure to consult the [codebook for the original data set](https://access.gesis.org/dbk/67378). Remember that only a subset of these variables (i.e., the numeric ones) are included in the synthetic data file that we use in this course (and that all values < 0 have been recoded as `NA`).
```

```{block, box.title = "Exercise 5", box.body = list(fill = "white"), box.icon = "fa-star"}
To further increase reproducibility, let's add some information about your `R` session (OS, `R` version, loaded packages) to the document. Add a section named *Reproducibility information* at the end of your document and display this information there.
```

```{block, opts.label = "clues", box.icon = "fa-lightbulb"}
You can use a `base R` function to access information about your session (function name hint 1) and another one for printing it (function name hint 2). You can exclude information about your `locale` settings there.
```

```{r session-info, eval=FALSE, box.title = "Solution 5", solution = TRUE, box.icon = "fa-check"}
sessionInfo() %>% 
  print(locale = FALSE)
```
