---
title: "Tools and Workflows for Reproducible Research in the Quantitative Social Sciences"
subtitle: "Using Git and GitHub"
author: "Bernd Weiß"
date: "2022-11-18"
presenter: Bernd
editor_options: 
  chunk_output_type: console
---
layout: true


```{r child = "content/config/sessions_setup.Rmd"}
```
---

class: center, middle
# Overview

---


## Recap

- Basic introduction to Git

- Git workflow

- Important Git concepts/commands (`git init`, `git add`, `git commit`, `git
log`, `git show`, `git clone`, `git checkout`, `git branch`, ...)

- ...

---


class: center, middle
# Please, let me in (authentication)

---



## Local or remote & HTTPS or SSH?

- A Git project is stored in a repository, which can be local or remote

- When using Git to access a remote repository (for backup or collaborative
work) on a remote server, you need to authenticate yourself to the server

- There are two ways of authentication: HTTPS or SSH

---


## Choose a remote server

- GitHub (which will be using)

- GitLab

- ...

---

## GitHub: Using personal access tokens

- These days, [authentication via personal access tokens (PAT) (and
https)](https://happygitwithr.com/https-pat.html) seems the way to go when using
GitHub

- In the following, I will illustrate the process using multiple screenshots

- Note that my explanation does not include any R/RStudio-related processes.
Johannes will talk about these things in more detail

---


## 

In GitHub, go to the `Settings` website:

```{r, out.width = "55%", echo = F}
woRkshoptools::include_picture("./f_github-pat-0-settings.png")
```

---


## 

Next, go to the `Developer Settings` entry: 

```{r, out.width = "90%", echo = F}
woRkshoptools::include_picture("./f_github-pat-1-settings.png")
```

---


## 

Then, choose `Tokens (classic)`:

```{r, out.width = "90%", echo = F}
woRkshoptools::include_picture("./f_github-pat-2-developer.png")
```

---


## 

And, generate a new token; important, save this Token (e.g.,
`ghp_FqamoGsx6PcEnpwnXyh...`):

```{r, out.width = "90%", echo = F}
woRkshoptools::include_picture("./f_github-pat-3-create-token.png")
```

---


## 

When you are now cloning a new repository (or pushing/pulling for the first
time), you will be asked once to enter your username...

```{r, out.width = "90%", echo = F}
woRkshoptools::include_picture("./f_github-pat-4-enter-credent.png")
```

---


## 

... and your Token (even though it says "Password"):

```{r, out.width = "90%", echo = F}
woRkshoptools::include_picture("./f_github-pat-4-enter-credent2.png")
```

---


## 

If, for whatever reason, you decide to reset/remove your credentials, you can do so
using the [Windows Credentials
Manager](https://support.microsoft.com/en-us/windows/accessing-credential-manager-1b5c916a-6a16-889f-8581-fc16e8165ac0)
(in German: "Anmeldeinformationsverwaltung")


```{r, out.width = "70%", echo = F}
woRkshoptools::include_picture("./f_github-pat-5-wincred.png")
```

---


## Setting up SSH

- SSH is a network protocol that comes in handy, when you work with remote
repositories and when you do not want to type-in your passwort every time you
pull (fetch) or push (send) from a remote repository. You still need to
authenticate yourself, though.

- To work with Git on you local computer, you do not need SSH (= Secure Shell). 

- Authentication in SSH (which is also the name of the program) works by using a
private and a public key (usually the public key has the file extension `.pub`,
e.g., my public key is `id_rsa.pub`). When you start working with SSH for the
very first time, you have to create both keys.

---


## 

- The private key remains on your local computer and you have to make sure that
it is safe -- it is a simple text file and it is your password now, and everyone
who has your private key can access your files. Again, everyone who has your
private key has your password!

This is what my *public key* looks like:

    ssh-rsa
    AAAAB3NzaC1yc2EAAAABIwAAAQEAyOQ9RT6TkfgkdO2NspzdVJE5CZ03yYAhVwLGo
    CrI3E9/Ix0MAySunXExjhsQi2XkhPBjLOEahYuuLaAWHuBc7apUPRNSBy+mdUHnH3
    0BdTQijQ6vj3RL99HO4yrZnipIlkS5ufw/+hpbXXOzSOqTvyGtL9ygm3eA2HDSQtz
    2ptFq8anODJDKrgTbNLb/YZ9KDIcpdO/Sfk4LtvaGF3tIFlyE+pogNmN4eWiYg9Xv
    25BhVVxWMHadRFLeDastWO4SedriEHzQYaNgxVNTufqolJ0nbg4R//fVDxjR2SbzV
    AHLZ+eVPUx+vzcPVMP9wYPcnii9YLiSRy+hlUAOR/kXeQ== berndweiss

---


class: middle
## 

**The *public key* (not the private key!) has to be stored at the
GitHub/GitLab/... website. Now, everyone who has your public key can encrypt
files (that are sent to you via the internet) but only you (or anyone else who
has your private key) can decrypt the files. And, for that reasons you do not
have to login everytime you push/pull files from the remote repository.**

---


## 

- How to setup SSH on you computer is explained on this website:
https://docs.gitlab.com/ce/ssh/README.html ("Generate an SSH key pair")

- The most important point is that `ssh` is able to find your key pair, i.e., it
needs to be located in your HOME folder

If everything works well, you should receive the following friendly welcome
message after typing in the command `ssh -T git@github.com`:

> Hi berndweiss! You've successfully authenticated, but GitHub does not provide shell access.

---


## Exercise

- In a previous exercise (see [](git:exercise1)), you have created your own repository (let's call it `your-new-repo`)

- Now, go to your GitHub account and create a new repository on GitHub 
    - Startpage -> tab "Repositories" -> green button "New"
    - Enter a new "Repository name"
    - Make it "Private" (unless you have something important to share)
    - Do **not** check any of the "Initialize this repository with" boxes
    - Hit the green "Create repository" button
    - Choose SSH or HTTPS as protocol ("Quick setup — if you’ve done this kind of thing before")
    - Look for "...or push an existing repository from the command line"
    - SSH:
        - Copy the line `git remote add origin git@github.com:berndweiss/your-repo-name.git` 
        - Execute the command `git remote add origin git@github.com:berndweiss/your-repo-name.git` in the Git Bash in your local repository (`your-new-repo`)
    - HTTPS:
        - Copy the line `git remote add origin https://github.com/berndweiss/your-repo-name.git` 
        - Execute the command `git remote add origin https://github.com/berndweiss/your-repo-name.git` in the Git Bash in your local repository (`your-new-repo`)
    - Make sure that `git status` shows a clean repository
    - Now you can run your first `git push origin main`
    - Reload the GitHub page via F5; you now should see the content of your local repo `your-new-repo`
    - ...
    - You can delete a GitHub repository via the tab "Settings" -> "Options", then scroll down -> "Danger Zone" -> "Delete this repository"

---



